---
title: "DeepDream - Subject 2 Physiological Analysis"
author:    |
    | Colin Ayres  
    | c.ayres@student.uw.edu.pl
date: "January 2025"
output: html_document
chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Packages and Load Data

```{r warning=FALSE, message=FALSE}
# Load necessary packages
library(tidyverse)
library(here)
library(ggplot2)
library(zoo)
library(patchwork)
library(mgcv)

# Set working directory
mds_dir <- here::here()

# Import data
data <- read.csv("pilot/2_DeviceData.csv", sep=";")

```

## Splitting Vectors into Directional Components

```{r}
# Splitting CombinedEyeDirection column into directional vectors
values <- gsub("[()]", "", data$CombinedEyeDirection) %>%
  strsplit(",") %>%
  lapply(as.numeric)
data$CombinedEyeDirectionX <- sapply(values, `[`, 1)
data$CombinedEyeDirectionY <- sapply(values, `[`, 2)
data$CombinedEyeDirectionZ <- sapply(values, `[`, 3)
```

## Interpolate Values During Blinking

While blinking, pupil values are recorded as negative. I replace negative values with interpolated values below. For example, (5, -3, -2, -4, 9) will become (5, 6, 7, 8, 9).

```{r}
# Interpolate blinking values
data_interpolated <- data %>%
  # Replace negative values with NA
  mutate(
    LeftEyePupilSize = ifelse(LeftEyePupilSize < 0, NA, LeftEyePupilSize),
    RightEyePupilSize = ifelse(RightEyePupilSize < 0, NA, RightEyePupilSize),
    CombinedEyeDirectionX = ifelse(CombinedEyeDirectionX < -0.98, NA, CombinedEyeDirectionX),
    CombinedEyeDirectionY = ifelse(CombinedEyeDirectionY < -0.98, NA, CombinedEyeDirectionY),
    CombinedEyeDirectionZ = ifelse(CombinedEyeDirectionZ < -0.98, NA, CombinedEyeDirectionZ)
  ) %>%
  # Linearly interpolate missing (NA) values
  mutate(
    LeftEyePupilSize = zoo::na.approx(LeftEyePupilSize, na.rm = FALSE),
    RightEyePupilSize = zoo::na.approx(RightEyePupilSize, na.rm = FALSE),
    CombinedEyeDirectionX = zoo::na.approx(CombinedEyeDirectionX, na.rm = FALSE),
    CombinedEyeDirectionY = zoo::na.approx(CombinedEyeDirectionY, na.rm = FALSE),
    CombinedEyeDirectionZ = zoo::na.approx(CombinedEyeDirectionZ, na.rm = FALSE)

  ) %>%
  # Fill any remaining NAs at the very start or very beginning of the dataset
  mutate(
    LeftEyePupilSize = zoo::na.locf(LeftEyePupilSize, na.rm = FALSE, fromLast = TRUE),
    LeftEyePupilSize = zoo::na.locf(LeftEyePupilSize, na.rm = FALSE),
    RightEyePupilSize = zoo::na.locf(RightEyePupilSize, na.rm = FALSE, fromLast = TRUE),
    RightEyePupilSize = zoo::na.locf(RightEyePupilSize, na.rm = FALSE),
    CombinedEyeDirectionX = zoo::na.locf(CombinedEyeDirectionX, na.rm = FALSE, fromLast = TRUE),
    CombinedEyeDirectionX = zoo::na.locf(CombinedEyeDirectionX, na.rm = FALSE),
    CombinedEyeDirectionY = zoo::na.locf(CombinedEyeDirectionY, na.rm = FALSE, fromLast = TRUE),
    CombinedEyeDirectionY = zoo::na.locf(CombinedEyeDirectionY, na.rm = FALSE),
    CombinedEyeDirectionZ = zoo::na.locf(CombinedEyeDirectionZ, na.rm = FALSE, fromLast = TRUE),
    CombinedEyeDirectionZ = zoo::na.locf(CombinedEyeDirectionZ, na.rm = FALSE)
    )

# Getting rid of "Countdown" (It takes 2 seconds and only appears in one of the conditions, messes up legend)
# Also getting rid of values for which Luminosity = 0
data_interpolated <- data_interpolated %>%
  mutate(Phase = ifelse(Phase == "Countdown", "BeforeRecording", Phase)) %>%
  filter(Luminosity != 0)

# Uncomment to visually check first 190 data points to make sure interpolation is correct 
# head(data$LeftEyePupilSize, 190)
# head(data_interpolated$LeftEyePupilSize, 190)
```

## Plotting Pupil Size

I plot pupil size against time, using the phase of the experiment as background for the chart.

```{r, echo=FALSE}
# Plot of LeftEyePupilSize for Subject 2, Interpolated Negative Values
ggplot(data_interpolated, aes(x = ExperimentSecond, y = LeftEyePupilSize)) +
  # Add background rectangles for phases
  geom_rect(aes(xmin = ExperimentSecond, xmax = lead(ExperimentSecond, default = max(ExperimentSecond)),
                ymin = -Inf, ymax = Inf, fill = Phase),
            color = NA, alpha = 0.3) +
  # Add black line for pupil size
  geom_line(color = "red") +
  theme_minimal() +
  labs(
    title = "Subject 2 Left Pupil Size During DayDream Experiment",
    x = "Time (seconds)",
    y = "Left Pupil Size",
    fill = "Phase"
  ) +
  # Adjust fill colors for phases
  scale_fill_manual(values = c(
    "BeforeRecording" = "blue",
    "Countdown" = "lightgreen",
    "Recording" = "pink",
    "SegmentFinished" = "yellow",
    "UIFeedback" = "purple",
    "Video" = "orange"
  )) +
  theme(legend.position = "bottom")

# Plot of RightEyePupilSize for Subject 2, Interpolated Negative Values
ggplot(data_interpolated, aes(x = ExperimentSecond, y = RightEyePupilSize)) +
  # Add background rectangles for phases
  geom_rect(aes(xmin = ExperimentSecond, xmax = lead(ExperimentSecond, default = max(ExperimentSecond)),
                ymin = -Inf, ymax = Inf, fill = Phase),
            color = NA, alpha = 0.3) +
  # Add black line for pupil size
  geom_line(color = "black") +
  theme_minimal() +
  labs(
    title = "Subject 2 Right Pupil Size During DayDream Experiment",
    x = "Time (seconds)",
    y = "Right Pupil Size",
    fill = "Phase"
  ) +
  # Adjust fill colors for phases
  scale_fill_manual(values = c(
    "BeforeRecording" = "blue",
    "Countdown" = "lightgreen",
    "Recording" = "pink",
    "SegmentFinished" = "yellow",
    "UIFeedback" = "purple",
    "Video" = "orange"
  )) +
  theme(legend.position = "bottom")

# Reshape the data to long format for both pupil sizes
data_long <- data_interpolated %>%
  select(ExperimentSecond, Phase, LeftEyePupilSize, RightEyePupilSize) %>%
  pivot_longer(
    cols = c(LeftEyePupilSize, RightEyePupilSize),
    names_to = "Eye",
    values_to = "PupilSize"
  )

# Plot both LeftEyePupilSize and RightEyePupilSize on the same chart
ggplot(data_long, aes(x = ExperimentSecond, y = PupilSize, color = Eye)) +
  # Add background rectangles for phases
  geom_rect(aes(xmin = ExperimentSecond, 
                xmax = lead(ExperimentSecond, default = max(ExperimentSecond)),
                ymin = -Inf, ymax = Inf, fill = Phase),
            inherit.aes = FALSE, color = NA, alpha = 0.3) +
  # Add lines for pupil sizes
  geom_line(linewidth = 1) +
  theme_minimal() +
  labs(
    title = "Subject 2 Pupil Sizes During DayDream Experiment",
    x = "Time (seconds)",
    y = "Pupil Size",
    fill = "Phase",
    color = "Eye"
  ) +
  # Adjust fill colors for phases
  scale_fill_manual(values = c(
    "BeforeRecording" = "blue",
    "Countdown" = "lightgreen",
    "Recording" = "pink",
    "SegmentFinished" = "yellow",
    "UIFeedback" = "purple",
    "Video" = "orange"
  )) +
  # Set line colors for Left and Right pupil sizes
  scale_color_manual(values = c(
    "LeftEyePupilSize" =  adjustcolor( "red", alpha.f = 0.5),
    "RightEyePupilSize" =  adjustcolor( "black", alpha.f = 0.5)
  )) +
  theme(legend.position = "bottom")
```

## Plotting Pupil Size by Video Mode (Normal/Hallucination) and Visual Type (Monoscopic/Sterescopic)

```{r warning=FALSE, message=FALSE}
# Define a function to create a single pupil/luminosity plot
create_pupil_plot <- function(data, title, y_limits, luminosity_factor) {
  ggplot(data, aes(x = ExperimentSecond - ExperimentSecond[1])) +
    geom_smooth(aes(y = round(LeftEyePupilSize, 3), color = "Left Pupil Size"), se = FALSE, method = "loess", span = 0.012) +
    geom_smooth(aes(y = round(RightEyePupilSize, 3), color = "Right Pupil Size"), se = FALSE, method = "loess", span = 0.012) +
    geom_line(aes(y = Luminosity * luminosity_factor, color = "Luminosity")) +
    scale_y_continuous(
      name = "Pupil Size (mm)",
      limits = y_limits,
      sec.axis = sec_axis(~ . / luminosity_factor, name = "Luminosity")
    ) +
    theme_minimal() +
    labs(
      title = title,
      x = "Time (seconds)",
      y = "Pupil Size (mm)",
      color = "Legend"
    ) +
    scale_color_manual(
      values = c(
        "Left Pupil Size" = "red",
        "Right Pupil Size" = "black",
        "Luminosity" = adjustcolor("green", alpha.f = 0.75)
      ),
      breaks = c("Left Pupil Size", "Right Pupil Size", "Luminosity")
    )
}

# Define common y-axis limits
common_y_limits <- c(1, 6)  # Adjust based on your data
luminosity_factor <- 14     # Scaling factor for Luminosity

# Create individual plots for the four conditions
plot1 <- create_pupil_plot(
  data_interpolated %>% 
    filter(data_interpolated$VideoMode == "Normal", data_interpolated$VideoName == "Main_Mono_HighLayer_Strong", data_interpolated$Phase == "Video"), "Normal Mono", common_y_limits, luminosity_factor)
plot2 <- create_pupil_plot(
  data_interpolated %>% 
    filter(data_interpolated$VideoMode == "Normal", data_interpolated$VideoName == "Main_Stereo_HighLayerStrong", data_interpolated$Phase == "Video"), "Normal Stereo", common_y_limits, luminosity_factor)
plot3 <- create_pupil_plot(
  data_interpolated %>% 
    filter(data_interpolated$VideoMode == "Hallucinations", data_interpolated$VideoName == "Main_Mono_HighLayer_Strong", data_interpolated$Phase == "Video"), "Hallucination Mono", common_y_limits, luminosity_factor)
plot4 <- create_pupil_plot(
  data_interpolated %>% 
    filter(data_interpolated$VideoMode == "Hallucinations", data_interpolated$VideoName == "Main_Stereo_HighLayerStrong", data_interpolated$Phase == "Video"), "Hallucination Stereo", common_y_limits, luminosity_factor)

# Combine all plots into a 2x2 layout and display
p <- (plot1 + plot2) / (plot3 + plot4) +
  plot_layout(guides = "collect") + # Collect the legends into one
  plot_annotation(title = "Subject 2 Pupil Size and Luminosity During DayDream Video") # Add a supertitle

print(p)
print(plot1)
print(plot2)
print(plot3)
print(plot4)
```

## Normalizing Pupil Size by Luminosity

```{r warning=FALSE, message=FALSE}

# Fit a linear regression model for pupil size based on luminosity; take average of left and right pupil sizes
fit_model <- function(data) {
  lm((LeftEyePupilSize + RightEyePupilSize) / 2 ~ Luminosity, data)
}

# Add normalized pupil size to the data, look at residuals
normalize_pupil_size <- function(data, model) {
  data <- data %>%
    mutate(
      AvgPupilSize = (LeftEyePupilSize + RightEyePupilSize) / 2,
      PredictedPupilSize = predict(model, newdata = data),
      ResidualPupilSize = AvgPupilSize - PredictedPupilSize
    )
  return(data)
}

# Normalized pupil size plotting
create_pupil_plot <- function(data, title, y_limits, luminosity_factor) {
  ggplot(data, aes(x = ExperimentSecond - ExperimentSecond[1])) +
      geom_smooth(aes(y = round(LeftEyePupilSize, 3), color = "Left Pupil Size"), se = FALSE, method = "loess", span = 0.012) +
    geom_smooth(aes(y = round(RightEyePupilSize, 3), color = "Right Pupil Size"), se = FALSE, method = "loess", span = 0.012) +
    geom_line(aes(y = Luminosity * luminosity_factor, color = "Luminosity")) +
geom_smooth(aes(y = ResidualPupilSize, color = "Residual Pupil Size"), se = FALSE, method = "loess", span = 0.012) +
    scale_y_continuous(
      name = "Pupil Size (mm)",
      limits = y_limits,
      sec.axis = sec_axis(~ . / luminosity_factor, name = "Luminosity")
    ) +
    theme_minimal() +
    labs(
      title = title,
      x = "Time (seconds)",
      y = "Pupil Size (mm)",
      color = "Legend"
    ) +
    scale_color_manual(
      values = c(
        "Left Pupil Size" = "red",
        "Right Pupil Size" = "black",
        "Luminosity" = adjustcolor("green", alpha.f = 0.75),
        "Residual Pupil Size" = "blue"
      ),
      breaks = c("Left Pupil Size", "Right Pupil Size", "Luminosity", "Residual Pupil Size")
    )
}

# Define common y-axis limits
common_y_limits <- c(-2, 5)  # Adjust based on your data
luminosity_factor <- 14     # Scaling factor for Luminosity

# Filter data to only include experiment
filtered_data <- data_interpolated %>%
  filter(
    VideoMode %in% c("Normal", "Hallucinations") &
    VideoName %in% c("Main_Mono_HighLayer_Strong", "Main_Stereo_HighLayerStrong") &
    Phase == "Video"
  )

filtered_data <- filtered_data %>%
  mutate(InverseLuminosity = 1 / Luminosity)

# Linear model
model_linear <- lm((LeftEyePupilSize + RightEyePupilSize) / 2 ~ InverseLuminosity, data = filtered_data)

# Non-linear(generalized additive) model
model_gam <- gam((LeftEyePupilSize + RightEyePupilSize) / 2 ~ s(InverseLuminosity, k = 10), data = filtered_data)


# Normalization
data_linear <- normalize_pupil_size(filtered_data, model_linear)
data_gam <-normalize_pupil_size(filtered_data, model_gam)

# Create plots with residuals from linear model
plot1 <- create_pupil_plot(
  data_linear %>% 
    filter(data_linear$VideoMode == "Normal", data_linear$VideoName == "Main_Mono_HighLayer_Strong", data_linear$Phase == "Video"), "Normal Mono Linear", common_y_limits, luminosity_factor)
plot2 <- create_pupil_plot(
  data_linear %>% 
    filter(data_linear$VideoMode == "Normal", data_linear$VideoName == "Main_Stereo_HighLayerStrong", data_linear$Phase == "Video"), "Normal Stereo Linear", common_y_limits, luminosity_factor)
plot3 <- create_pupil_plot(
  data_linear %>% 
    filter(data_linear$VideoMode == "Hallucinations", data_linear$VideoName == "Main_Mono_HighLayer_Strong", data_linear$Phase == "Video"), "Hallucination Mono Linear", common_y_limits, luminosity_factor)
plot4 <- create_pupil_plot(
  data_linear %>% 
    filter(data_linear$VideoMode == "Hallucinations", data_linear$VideoName == "Main_Stereo_HighLayerStrong", data_linear$Phase == "Video"), "Hallucination Stereo Linear ", common_y_limits, luminosity_factor)

# Combine all plots into a 2x2 layout and display
p <- (plot1 + plot2) / (plot3 + plot4) +
  plot_layout(guides = "collect") + # Collect the legends into one
  plot_annotation(title = "Subject 2 Pupil Size Normalized by Luminosity, Linear Model") # Add a supertitle

print(p)
print(plot1)
print(plot2)
print(plot3)
print(plot4)

# Create plots with residuals from non-linear (generalized additive) model
plot1 <- create_pupil_plot(
  data_gam %>% 
    filter(data_gam$VideoMode == "Normal", data_gam$VideoName == "Main_Mono_HighLayer_Strong", data_gam$Phase == "Video"), "Normal Mono GAM, k = 10", common_y_limits, luminosity_factor)
plot2 <- create_pupil_plot(
  data_gam %>% 
    filter(data_gam$VideoMode == "Normal", data_gam$VideoName == "Main_Stereo_HighLayerStrong", data_gam$Phase == "Video"), "Normal Stereo GAM", common_y_limits, luminosity_factor)
plot3 <- create_pupil_plot(
  data_gam %>% 
    filter(data_gam$VideoMode == "Hallucinations", data_gam$VideoName == "Main_Mono_HighLayer_Strong", data_gam$Phase == "Video"), "Hallucination Mono GAM", common_y_limits, luminosity_factor)
plot4 <- create_pupil_plot(
  data_gam %>% 
    filter(data_gam$VideoMode == "Hallucinations", data_gam$VideoName == "Main_Stereo_HighLayerStrong", data_gam$Phase == "Video"), "Hallucination Stereo GAM", common_y_limits, luminosity_factor)

# Combine all plots into a 2x2 layout and display
p <- (plot1 + plot2) / (plot3 + plot4) +
  plot_layout(guides = "collect") + # Collect the legends into one
  plot_annotation(title = "Subject 2 Pupil Size Normalized by Luminosity, Generalized Additive Model") # Add a supertitle

print(p)
print(plot1)
print(plot2)
print(plot3)
print(plot4)

#summary(model)
#summary(model_gam)

```

## Plotting Gaze Shifts Alongside Pupil Diameter

```{r warning=FALSE, message=FALSE}
# Function to find where gaze is relatively stable
find_stable_gaze_regions <- function(data, threshold, duration, cooldown) {
  stable_regions <- data.frame(StartTime = numeric(), EndTime = numeric())
  in_region <- FALSE
  start_time <- NA
  last_end_time <- -Inf 

  for (i in 1:nrow(data)) {
    if (in_region) {
      # Check if the gaze is still stable
      if (abs(data$CombinedEyeDirectionX[i] - data$CombinedEyeDirectionX[start_time]) <= threshold &&
          abs(data$CombinedEyeDirectionY[i] - data$CombinedEyeDirectionY[start_time]) <= threshold) {
        if ((data$ExperimentSecond[i] - data$ExperimentSecond[start_time]) >= duration) {
          end_time <- i
          in_region <- FALSE
          stable_regions <- rbind(
            stable_regions,
            data.frame(
              StartTime = data$ExperimentSecond[start_time],
              EndTime = data$ExperimentSecond[end_time]
            )
          )
          last_end_time <- data$ExperimentSecond[end_time] + cooldown
        }
      } else {
        # Exit region if gaze is no longer stable
        in_region <- FALSE
      }
    } else {
      # Start new stable region if cooldown period has passed
      if (data$ExperimentSecond[i] > last_end_time &&
          abs(data$CombinedEyeDirectionX[i] - mean(data$CombinedEyeDirectionX[i:min(nrow(data), i + duration - 1)])) <= threshold &&
          abs(data$CombinedEyeDirectionY[i] - mean(data$CombinedEyeDirectionY[i:min(nrow(data), i + duration - 1)])) <= threshold) {
        start_time <- i
        in_region <- TRUE
      }
    }
  }

  return(stable_regions)
}

# Plotting function
create_pupil_plot <- function(data, title, y_limits, direction_factor, threshold, duration, cooldown) {
  stable_regions <- find_stable_gaze_regions(data, threshold, duration, cooldown)
  
  ggplot(data, aes(x = ExperimentSecond - ExperimentSecond[1])) +
    geom_smooth(aes(y = round(LeftEyePupilSize, 3), color = "Left Pupil Size"), se = FALSE, method = "loess", span = 0.012) +
    geom_smooth(aes(y = round(RightEyePupilSize, 3), color = "Right Pupil Size"), se = FALSE, method = "loess", span = 0.012) +
    #geom_smooth(aes(y = CombinedEyeDirectionX * direction_factor, color = "Gaze X Component"), se = FALSE, method = "loess", span = 0.012) +
    #geom_smooth(aes(y = CombinedEyeDirectionY * direction_factor, color = "Gaze Y Component"), se = FALSE, method = "loess", span = 0.012) +
    geom_line(aes(y = CombinedEyeDirectionX * direction_factor, color = "Gaze X Component")) +
    geom_line(aes(y = CombinedEyeDirectionY * direction_factor, color = "Gaze Y Component")) +
    geom_rect(
      data = stable_regions,
      aes(
        xmin = StartTime - data$ExperimentSecond[1],
        xmax = EndTime - data$ExperimentSecond[1],
        ymin = y_limits[1],
        ymax = y_limits[2]
      ),
      inherit.aes = FALSE,
      fill = "green",
      alpha = 0.45
    ) +
    scale_y_continuous(
      name = "Pupil Size (mm)",
      limits = y_limits,
      sec.axis = sec_axis(~ . / direction_factor, name = "Gaze Direction")
    ) +
    theme_minimal() +
    labs(
      title = title,
      x = "Time (seconds)",
      y = "Pupil Size (mm)",
      color = "Legend"
    ) +
    scale_color_manual(
      values = c(
        "Left Pupil Size" = "red",
        "Right Pupil Size" = "black",
        "Gaze X Component" = adjustcolor("blue", alpha.f = 0.75),
        "Gaze Y Component" = adjustcolor("orange", alpha.f = 0.75)
      )
    )
}

# Define parameters
common_y_limits <- c(-3, 4)  # Adjust based on data
direction_factor <- 5        # Scaling factor for gaze direction
x_tolerance <- 0.25          # Gaze difference tolerance (adjust as needed)
duration_threshold <- 2     # Minimum number of points for stability (adjust as needed)
cooldown = 1               # Minimum number of points before another rectangle can be drawn (adjust as needed)

# Create individual plots for the four conditions
plot1 <- create_pupil_plot(
  data_interpolated %>% 
    filter(data_interpolated$VideoMode == "Normal", data_interpolated$VideoName == "Main_Mono_HighLayer_Strong", data_interpolated$Phase == "Video"),
  "Normal Mono", common_y_limits, direction_factor, x_tolerance, duration_threshold, cooldown
)

plot2 <- create_pupil_plot(
  data_interpolated %>% 
    filter(data_interpolated$VideoMode == "Normal", data_interpolated$VideoName == "Main_Stereo_HighLayerStrong", data_interpolated$Phase == "Video"),
  "Normal Stereo", common_y_limits, direction_factor, x_tolerance, duration_threshold, cooldown
)

plot3 <- create_pupil_plot(
  data_interpolated %>% 
    filter(data_interpolated$VideoMode == "Hallucinations", data_interpolated$VideoName == "Main_Mono_HighLayer_Strong", data_interpolated$Phase == "Video"),
  "Hallucination Mono", common_y_limits, direction_factor, x_tolerance, duration_threshold, cooldown
)

plot4 <- create_pupil_plot(
  data_interpolated %>% 
    filter(data_interpolated$VideoMode == "Hallucinations", data_interpolated$VideoName == "Main_Stereo_HighLayerStrong", data_interpolated$Phase == "Video"),
  "Hallucination Stereo", common_y_limits, direction_factor, x_tolerance, duration_threshold, cooldown
)

#Combine all plots into a 2x2 layout
p <- (plot1 + plot2) / (plot3 + plot4) +
  plot_layout(guides = "collect") + 
  plot_annotation(title = paste(
    "Subject 2 Pupil Size with Gaze Directions During DayDream Video;",
    "Stable Gaze Regions Highlighted; Tolerance =", x_tolerance,
    ", Duration =", duration_threshold,
    "sec)"),
    theme = theme(plot.title = element_text(size = 9, hjust = 0.5)) 
    )

print(p)

print(plot1)
print(plot2)
print(plot3)
print(plot4)

```
