---
title: "Study 01 - Subjects Combined Eye-Tracking Data Analysis"
author:    |
    | Colin Ayres  
    | c.ayres@student.uw.edu.pl
date: "April 2025"
output: html_document
chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Packages

```{r warning=FALSE, message=FALSE}
# Load necessary packages
library(tidyverse)
library(here)
library(ggplot2)
library(zoo)
library(dplyr)
library(patchwork)
library(mgcv)
library(ggpubr)
library(rstatix)

# Set working directory
mds_dir <- here::here()

```

## Defining Functions to Clean & Combine Data

```{r}
# Splitting Vectors into Directional Components
split_eye_direction <- function(data) {
  data %>%
    mutate(
      # Remove parentheses and split into components
      values = strsplit(gsub("[()]", "", CombinedEyeDirection), ","),
      # Extract numeric values for each dimension
      CombinedEyeDirectionX = sapply(values, function(x) as.numeric(x[1])),
      CombinedEyeDirectionY = sapply(values, function(x) as.numeric(x[2])),
      CombinedEyeDirectionZ = sapply(values, function(x) as.numeric(x[3])),
      # Remove temporary values column
      .keep = "unused"
    )
}

# While blinking, pupil values are recorded as negative. I replace negative values with interpolated values below. For example, (5, -3, -2, -4, 9) will become (5, 6, 7, 8, 9).
# Interpolate blinking values
interpolate_blinking <- function(data) {
  data %>%
    # Replace invalid values with NA
    mutate(
      LeftEyePupilSize = ifelse(LeftEyePupilSize < 0, NA, LeftEyePupilSize),
      RightEyePupilSize = ifelse(RightEyePupilSize < 0, NA, RightEyePupilSize),
      CombinedEyeDirectionX = ifelse(CombinedEyeDirectionX < -0.98, NA, CombinedEyeDirectionX),
      CombinedEyeDirectionY = ifelse(CombinedEyeDirectionY < -0.98, NA, CombinedEyeDirectionY),
      CombinedEyeDirectionZ = ifelse(CombinedEyeDirectionZ < -0.98, NA, CombinedEyeDirectionZ)
    ) %>%
    # Linear interpolation for missing values
    mutate(
      across(c(LeftEyePupilSize, RightEyePupilSize, 
               CombinedEyeDirectionX, CombinedEyeDirectionY, CombinedEyeDirectionZ),
             ~ zoo::na.approx(., na.rm = FALSE))
    ) %>%
    # Fill leading/trailing NAs using last observation carried forward/backward
    mutate(
      LeftEyePupilSize = zoo::na.locf(zoo::na.locf(LeftEyePupilSize, na.rm = FALSE, fromLast = TRUE), na.rm = FALSE),
      RightEyePupilSize = zoo::na.locf(zoo::na.locf(RightEyePupilSize, na.rm = FALSE, fromLast = TRUE), na.rm = FALSE),
      CombinedEyeDirectionX = zoo::na.locf(zoo::na.locf(CombinedEyeDirectionX, na.rm = FALSE, fromLast = TRUE), na.rm = FALSE),
      CombinedEyeDirectionY = zoo::na.locf(zoo::na.locf(CombinedEyeDirectionY, na.rm = FALSE, fromLast = TRUE), na.rm = FALSE),
      CombinedEyeDirectionZ = zoo::na.locf(zoo::na.locf(CombinedEyeDirectionZ, na.rm = FALSE, fromLast = TRUE), na.rm = FALSE)
    )
}

# Getting rid of "Countdown" (It takes 2 seconds and only appears in one of the conditions, messes up legend)
# Also getting rid of values for which Luminosity = 0
clean_phase_luminosity <- function(data) {
  # Perform cleaning operations
  data %>%
    mutate(Phase = ifelse(Phase == "Countdown", "BeforeRecording", Phase)) %>%
    filter(Luminosity != 0)
}


clean_video_data <- function(df) {
  df %>%
    mutate(
      Condition = VideoName,
      Mode = case_when(
        grepl("HighLayer_Weak", VideoName) ~ "Hallucination",
        grepl("Normal", VideoName) ~ "Normal",
        grepl("LowLayer_Strong", VideoName) ~ "Hallucination",
        grepl("MidLayer_Strong", VideoName) ~ "Hallucination",
        grepl("MidLayer_Weak", VideoName) ~ "Hallucination",
        grepl("HighLayer_Strong", VideoName) ~ "Hallucination",
        grepl("LowLayer_Weak", VideoName) ~ "Hallucination",
        TRUE ~ as.character(VideoMode)
      ),
      Type = case_when(
        grepl("HighLayer_Weak", VideoName) ~ "HighLayer",
        grepl("Normal", VideoName) ~ "Normal",
        grepl("LowLayer_Strong", VideoName) ~ "LowLayer",
        grepl("MidLayer_Strong", VideoName) ~ "MidLayer",
        grepl("MidLayer_Weak", VideoName) ~ "MidLayer",
        grepl("HighLayer_Strong", VideoName) ~ "HighLayer",
        grepl("LowLayer_Weak", VideoName) ~ "LowLayer",
        TRUE ~ as.character(VideoMode)
      ),
      Strength = case_when(
        grepl("HighLayer_Weak", VideoName) ~ "Weak",
        grepl("Normal", VideoName) ~ "Normal",
        grepl("LowLayer_Strong", VideoName) ~ "Strong",
        grepl("MidLayer_Strong", VideoName) ~ "Strong",
        grepl("MidLayer_Weak", VideoName) ~ "Weak",
        grepl("HighLayer_Strong", VideoName) ~ "Strong",
        grepl("LowLayer_Weak", VideoName) ~ "Weak",
        TRUE ~ as.character(VideoMode)
      )
    ) %>%
    select(-VideoName, -VideoMode)  # Drop VideoName and VideoMode
}

```

## Load and Process All Subject Data

```{r warning=FALSE, message=FALSE}
import_subjects <- function(data_dir = "Study_01") {
  # Get list of all device data files
  file_list <- list.files(
    path = here(data_dir),
    pattern = "\\d+_DeviceData\\.csv",
    full.names = TRUE
  )
  
  # Read and combine files
  map_df(file_list, ~ {
    # Read the file as raw text to get the header
    raw_lines <- read_lines(.x)
    header <- raw_lines[1]
    # Remove trailing semicolon from header
    header_clean <- sub(";+$", "", header)
    col_names <- unlist(strsplit(header_clean, ";"))
    
    # Read the data, skipping the header row
    read_delim(.x, 
               delim = ";",
               col_names = col_names,  # Use cleaned column names
               skip = 1,  # Skip the header row
                col_types = cols(
                 Id = col_character(),
                 VideoName = col_character(),
                 VideoMode = col_character(),
                 Phase = col_character(),
                 DevicePosition = col_character(),
                 DeviceRotation = col_character(),
                 CombinedEyeOrigin = col_character(),
                 LeftEyeOrigin = col_character(),
                 LeftEyeDirection = col_character(),
                 LeftEyePositionInSensorArea = col_character(),
                 RightEyeOrigin = col_character(),
                 RightEyeDirection = col_character(),
                 RightEyePositionInSensorArea = col_character(),
                 CombinedEyeDirection = col_character(),
                 .default = col_double()  # All other columns as numeric
               ),
               show_col_types = FALSE)
  }) %>%
    arrange(Id)  # Sort by Id (as character)
}

# Load the data into a dataframe
#device_data <- import_subjects()

# View the first few rows
#head(device_data)

# Check column names
#names(device_data)

# Process with subject isolation
process_all_subjects <- function(raw_data) {
  raw_data %>%
    group_by(Id) %>%
    group_modify(~ {
      .x %>%
        clean_phase_luminosity() %>%
        split_eye_direction() %>%
        interpolate_blinking() %>%
        clean_video_data()
    }) %>%
    ungroup() %>%
    arrange(Id, ExperimentSecond)
}

# Import/process data
processed_device_combined <- import_subjects() %>% process_all_subjects()

```


## Plotting Functions

```{r warning=FALSE, message=FALSE}
# Add normalized pupil size to the data, look at residuals
normalize_pupil_size <- function(data, model) {
  data <- data %>%
    mutate(
      AvgPupilSize = (LeftEyePupilSize + RightEyePupilSize) / 2,
      PredictedPupilSize = predict(model, newdata = data),
      ResidualPupilSize = AvgPupilSize - PredictedPupilSize
    )
  return(data)
}

# Normalized pupil size plotting
create_normalized_pupil_plot <- function(data, title, y_limits, luminosity_factor) {
  ggplot(data, aes(x = ExperimentSecond - ExperimentSecond[1])) +
      geom_smooth(aes(y = round(AvgPupilSize, 3), color = "Avg. Pupil Size"), se = FALSE, method = "loess", span = 0.012) +
    geom_line(aes(y = Luminosity * luminosity_factor, color = "Luminosity")) +
geom_smooth(aes(y = ResidualPupilSize, color = "Residual Pupil Size"), se = FALSE, method = "loess", span = 0.012) +
    scale_y_continuous(
      name = "Pupil Size (mm)",
      limits = y_limits,
      sec.axis = sec_axis(~ . / luminosity_factor, name = "Luminosity")
    ) +
    theme_minimal() +
    labs(
      title = title,
      x = "Time (seconds)",
      y = "Pupil Size (mm)",
      color = "Legend"
    ) +
    scale_color_manual(
      values = c(
        "Avg. Pupil Size" = "black",
        "Luminosity" = adjustcolor("green", alpha.f = 0.75),
        "Residual Pupil Size" = "blue"
      ),
      breaks = c("Avg. Pupil Size", "Luminosity", "Residual Pupil Size")
    )
}

# All data plotting
create_faceted_pupil_plots <- function(data, title, all_plot_y_limits, ncol = 5) {
  plot_data <- data %>%
    group_by(Id) %>%
    mutate(
      TimeElapsed = ExperimentSecond - min(ExperimentSecond),  # Relative time per subject
      AvgPupilSize = (LeftEyePupilSize + RightEyePupilSize) / 2
    ) %>%
    ungroup()
  
  ggplot(plot_data, aes(x = TimeElapsed, y = AvgPupilSize)) +
    geom_smooth(method = "loess", span = 0.015, se = FALSE, color = "#2c7bb6") +
    geom_point(size = 0.5, alpha = 0.2, color = "#636363") + 
    facet_wrap(~ Id, ncol = ncol, drop = FALSE) +
    theme_minimal(base_size = 12) +
    labs(
      title = title,
      x = "Time",
      y = "Average Pupil Size (mm)"
    ) +
    scale_y_continuous(
      limits = all_plot_y_limits,
      breaks = seq(2, 5, by = 0.5)
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      strip.background = element_rect(fill = "#f0f0f0"),
      strip.text = element_text(face = "bold", size = 10),
      panel.spacing = unit(1.2, "lines"),
      axis.text.x = element_blank() 
    )
}

```

## Modelling Pupil Size vs. Luminosity

```{r warning=FALSE, message=FALSE}
# Define common variables
common_y_limits <- c(-2, 5)
all_plot_y_limits <- c(2, 5)
luminosity_factor <- 14

# Filter data to only include experiment
filtered_data <- processed_device_combined %>%
  filter(
    Mode %in% c("Normal", "Hallucination"),
    Phase == "Video"
  )

# Add InverseLuminosity Column
filtered_data <- filtered_data %>%
  mutate(InverseLuminosity = 1 / Luminosity)

# Non-linear (generalized additive) model
model_gam <- gam((LeftEyePupilSize + RightEyePupilSize) / 2 ~ s(InverseLuminosity, k = 10), 
                 data = filtered_data[!is.na(filtered_data$InverseLuminosity), ])
data_gam <- normalize_pupil_size(filtered_data, model_gam)
data_gam <- data_gam %>% mutate(Id = factor(Id))

# Check unique Id and Condition values
unique_ids <- unique(data_gam$Id[!is.na(data_gam$Id)])
unique_conditions <- unique(data_gam$Condition[!is.na(data_gam$Condition)])
cat("Unique Ids:", unique_ids, "\n")
cat("Unique Conditions:", unique_conditions, "\n")

# Ensure exactly 7 conditions (for splitting into 4 and 3)
if (length(unique_conditions) != 7) {
  warning("Expected 7 unique Conditions, found ", length(unique_conditions))
}

# Create plots for each subject, arranged in 4x4 grids
for (id in unique_ids) {
  # Generate plots for all conditions for this subject
  plots <- list()
  for (cond in unique_conditions) {
    plot_data <- data_gam %>%
      filter(
        Id == id,
        Condition == cond,
        Phase == "Video",
        !is.na(ResidualPupilSize)
      )
    
    if (nrow(plot_data) > 0) {
      plot_title <- sprintf("%s", cond)
      p <- create_normalized_pupil_plot(
        data = plot_data,
        title = plot_title,
        y_limits = common_y_limits,
        luminosity_factor = luminosity_factor
      )
      plots[[cond]] <- p
    } else {
      # Create an empty plot if no data
      p <- ggplot() + theme_void() + labs(title = sprintf("%s (No Data)", cond))
      plots[[cond]] <- p
    }
  }
  
  # Split plots into two groups: first 4 and last 3
  plots_first <- plots[1:4]
  plots_second <- plots[5:7]
  
  # Create 4x4 grid for first 4 plots
  grid_first <- wrap_plots(plots_first, ncol = 2, nrow = 2) + 
    plot_layout(guides = "collect") +
    plot_annotation(
      title = sprintf("Subject %s Pupil Size Normalized by Luminosity, Part 1", id),
      theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
    )
  
  # Create 4x4 grid for last 3 plots (with empty spaces)
  grid_second <- wrap_plots(plots_second, ncol = 2, nrow = 2) + 
    plot_layout(guides = "collect") +
    plot_annotation(
      title = sprintf("Subject %s Pupil Size Normalized by Luminosity, Part 2", id),
      theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
    )
  
  # Print both grids
  print(grid_first)
  print(grid_second)
  
  # Optionally save to files
  #ggsave(sprintf("subject_%s_part1.png", id), grid_first, width = 12, height = 12)
  #ggsave(sprintf("subject_%s_part2.png", id), grid_second, width = 12, height = 12)
}

# Create faceted plots for all subjects per condition (unchanged)
plots_faceted <- list()
for (cond in unique_conditions) {
  plot_data <- data_gam %>%
    filter(
      Condition == cond,
      Phase == "Video",
      !is.na(AvgPupilSize)
    )
  
  if (nrow(plot_data) > 0) {
    plot_title <- sprintf("All Subjects, Condition: %s", cond)
    p <- create_faceted_pupil_plots(
      data = plot_data,
      title = plot_title,
      all_plot_y_limits = all_plot_y_limits,
      ncol = 5
    )
    plots_faceted[[cond]] <- p
  }
}

# Print faceted plots
for (p in plots_faceted) {
  print(p)
}

# Print model summary
summary(model_gam)
```

## Statistical Tests for Pupil Data

```{r warning=FALSE, message=FALSE}
# Load required libraries
library(dplyr)
library(ggplot2)
library(rstatix)  # For pairwise_t_test and stat_pvalue_manual

# Create summary data for statistical testing
condition_data <- data_gam %>%
  filter(Phase == "Video", !is.na(Condition)) %>%
  group_by(Id, Condition) %>%
  summarize(
    AvgPupilSize = mean((LeftEyePupilSize + RightEyePupilSize) / 2, na.rm = TRUE),
    ResidualPupilSize = mean(ResidualPupilSize, na.rm = TRUE),
    .groups = "drop"
  )

# Get all unique conditions and generate pairwise comparisons
unique_conditions <- unique(condition_data$Condition)
comparisons <- combn(unique_conditions, 2, simplify = FALSE)

# Perform paired t-tests with Bonferroni correction
paired_t_avgpupil <- condition_data %>%
  pairwise_t_test(
    AvgPupilSize ~ Condition,
    paired = TRUE,
    comparisons = comparisons,
    p.adjust.method = "bonferroni"
  ) %>%
  mutate(p.adj.signif = case_when(
    p.adj < 0.001 ~ "***",
    p.adj < 0.01 ~ "**",
    p.adj < 0.05 ~ "*",
    TRUE ~ "ns"
  ))

paired_t_residual <- condition_data %>%
  pairwise_t_test(
    ResidualPupilSize ~ Condition,
    paired = TRUE,
    comparisons = comparisons,
    p.adjust.method = "bonferroni"
  ) %>%
  mutate(p.adj.signif = case_when(
    p.adj < 0.001 ~ "***",
    p.adj < 0.01 ~ "**",
    p.adj < 0.05 ~ "*",
    TRUE ~ "ns"
  ))

# Identify significant comparisons (where p.adj.signif != "ns" for either metric)
sig_comparisons <- unique(rbind(
  paired_t_avgpupil %>% filter(p.adj.signif != "ns") %>% select(group1, group2),
  paired_t_residual %>% filter(p.adj.signif != "ns") %>% select(group1, group2)
)) %>%
  split(1:nrow(.))

# Print t-test results for reference
print("Paired t-tests for AvgPupilSize:")
print(paired_t_avgpupil)
print("Paired t-tests for ResidualPupilSize:")
print(paired_t_residual)

# Generate plots for significant condition pairs
for (i in seq_along(sig_comparisons)) {
  pair <- sig_comparisons[[i]]
  cond1 <- pair$group1
  cond2 <- pair$group2
  
  # Filter data for this pair
  pair_data <- condition_data %>%
    filter(Condition %in% c(cond1, cond2))
  
  # Filter t-test results for this pair
  p_avg <- paired_t_avgpupil %>%
    filter((group1 == cond1 & group2 == cond2) | (group1 == cond2 & group2 == cond1))
  p_res <- paired_t_residual %>%
    filter((group1 == cond1 & group2 == cond2) | (group1 == cond2 & group2 == cond1))
  
  # Prepare p-value labels for significant results
  p_avg_label <- p_avg %>%
    filter(p.adj.signif != "ns") %>%
    mutate(label = sprintf("p = %.4f %s", p.adj, p.adj.signif))
  p_res_label <- p_res %>%
    filter(p.adj.signif != "ns") %>%
    mutate(label = sprintf("p = %.4f %s", p.adj, p.adj.signif))
  
  # Plot AvgPupilSize
  avg_plot <- ggplot(pair_data, aes(x = Condition, y = AvgPupilSize)) +
    geom_boxplot(width = 0.4, outlier.shape = NA) +
    geom_line(aes(group = Id), color = "gray", alpha = 0.4) +
    geom_point(aes(color = Condition), size = 2, alpha = 0.8) +
    {if (nrow(p_avg_label) > 0) stat_pvalue_manual(
      p_avg_label,
      label = "label",
      y.position = max(pair_data$AvgPupilSize, na.rm = TRUE) * 1.1
    )} +
    scale_x_discrete(limits = c(cond1, cond2)) +
    labs(
      title = sprintf("Mean Pupil Size: %s vs %s", cond1, cond2),
      y = "Mean Pupil Size",
      x = "Condition"
    ) +
    theme_minimal() +
    theme(legend.position = "none", axis.text.x = element_text(size = 8))
  
  # Plot ResidualPupilSize
  res_plot <- ggplot(pair_data, aes(x = Condition, y = ResidualPupilSize)) +
    geom_boxplot(width = 0.4, outlier.shape = NA) +
    geom_line(aes(group = Id), color = "gray", alpha = 0.4) +
    geom_point(aes(color = Condition), size = 2, alpha = 0.8) +
    {if (nrow(p_res_label) > 0) stat_pvalue_manual(
      p_res_label,
      label = "label",
      y.position = max(pair_data$ResidualPupilSize, na.rm = TRUE) * 1.1
    )} +
    scale_x_discrete(limits = c(cond1, cond2)) +
    labs(
      title = sprintf("Mean Residual Pupil Size: %s vs %s", cond1, cond2),
      y = "Mean Residual Pupil Size",
      x = "Condition"
    ) +
    theme_minimal() +
    theme(legend.position = "none", axis.text.x = element_text(size = 8))
  
  # Print plots
  print(avg_plot)
  print(res_plot)
}
```

## Defining Functions for Gaze Analysis

```{r warning=FALSE, message=FALSE}
parse_vector_string <- function(vec_string) {
  # Handle list input (e.g., data_gam$values)
  if (is.list(vec_string)) {
    vec_string <- paste0("(", paste(trimws(unlist(vec_string)), collapse = ", "), ")")
  }
  
  # Skip if NA or empty
  if (is.na(vec_string) || vec_string == "") return(NA)
  
  # Remove all non-numeric characters except commas, dots, and minus signs
  clean_string <- gsub("[^0-9.,-]", "", vec_string)
  parts <- strsplit(clean_string, ",\\s*")[[1]]
  result <- as.numeric(parts)
  
  if (length(result) != 3 || any(is.na(result))) {
    warning("Invalid vector format: ", vec_string)
    return(NA)
  }
  return(result)
}

# Function to convert Euler angles (roll, pitch, yaw) to a rotation matrix
euler_to_matrix <- function(euler) {
  # Unpack Euler angles (in degrees, convert to radians)
  roll  <- euler[1] * pi / 180
  pitch <- euler[2] * pi / 180
  yaw   <- euler[3] * pi / 180
  
  # Rotation matrix about the x-axis
  R_x <- matrix(c(
    1,         0,          0,
    0, cos(roll), -sin(roll),
    0, sin(roll),  cos(roll)
  ), nrow = 3, byrow = TRUE)
  
  # Rotation matrix about the y-axis
  R_y <- matrix(c(
    cos(pitch), 0, sin(pitch),
    0,          1,         0,
    -sin(pitch), 0, cos(pitch)
  ), nrow = 3, byrow = TRUE)
  
  # Rotation matrix about the z-axis
  R_z <- matrix(c(
    cos(yaw), -sin(yaw), 0,
    sin(yaw),  cos(yaw), 0,
    0,              0,   1
  ), nrow = 3, byrow = TRUE)
  
  # Combine rotations: R = R_z(yaw) %*% R_y(pitch) %*% R_x(roll)
  rot_matrix <- R_z %*% R_y %*% R_x
  return(rot_matrix)
}

# Function to calculate world gaze vectors
calculate_world_gaze_vectors <- function(data) {
  n_samples <- nrow(data)
  world_vectors <- matrix(NA, nrow = n_samples, ncol = 3)
  
  for (i in 1:n_samples) {
    # Parse DeviceRotation
    rotation_string <- as.character(data[i, "DeviceRotation"])
    rotation <- parse_vector_string(rotation_string)
    
    if (length(rotation) != 3 || any(is.na(rotation))) {
      warning(sprintf("Invalid DeviceRotation at row %d: %s", i, rotation_string))
      next
    }
    
    # Get rotation matrix
    rot_matrix <- euler_to_matrix(rotation)
    
    # Parse CombinedEyeDirection
    eye_string <- as.character(data[i, "values"])
    eye_direction <- parse_vector_string(eye_string)
    
    if (length(eye_direction) != 3 || any(is.na(eye_direction))) {
      warning(sprintf("Invalid CombinedEyeDirection at row %d: %s", i, eye_string))
      next
    }
    
    # Transform eye direction to world space
    world_vectors[i, ] <- rot_matrix %*% matrix(eye_direction, nrow = 3)
  }
  
  return(world_vectors)
}

# Calculate angle between vectors
calculate_angular_diff <- function(v1, v2) {
  if (any(is.na(v1)) || any(is.na(v2))) {
    return(NA)
  }
  v1_norm <- v1 / sqrt(sum(v1^2))
  v2_norm <- v2 / sqrt(sum(v2^2))
  if (any(is.na(v1_norm)) || any(is.na(v2_norm))) {
    return(NA)
  }
  dot_prod <- sum(v1_norm * v2_norm)
  dot_prod <- max(min(dot_prod, 1), -1)
  angle_rad <- acos(dot_prod)
  return(angle_rad * 180 / pi)
}

# Calculate compensated gaze stability
calculate_gaze_stability <- function(data, window_size = 30) {
  world_vectors <- calculate_world_gaze_vectors(data)
  n_samples <- nrow(world_vectors)
  angular_differences <- numeric(n_samples - 1)
  
  # Calculate angular differences between consecutive frames
  for (i in 2:n_samples) {
    if (any(is.na(world_vectors[i, ])) || any(is.na(world_vectors[i-1, ]))) {
      angular_differences[i-1] <- NA
    } else {
      angular_differences[i-1] <- calculate_angular_diff(
        world_vectors[i, ],
        world_vectors[i-1, ]
      )
    }
  }
  
  # Calculate rolling stability scores
  n_windows <- max(1, length(angular_differences) - window_size + 1)
  stability_scores <- numeric(n_windows)
  for (i in 1:n_windows) {
    window <- angular_differences[i:(i + window_size - 1)]
    if (all(is.na(window))) {
      stability_scores[i] <- NA
    } else {
      stability_scores[i] <- mean(window, na.rm = TRUE)
    }
  }
  
  list(
    stability_scores = stability_scores,
    angular_differences = angular_differences
  )
}

# Detect stable fixations
detect_fixations <- function(angular_differences, max_angle = 1.0, min_duration = 10) {
  n_samples <- length(angular_differences)
  fixation_periods <- list()
  current_start <- NULL
  
  for (i in 1:n_samples) {
    if (is.na(angular_differences[i]) || angular_differences[i] <= max_angle) {
      if (is.null(current_start)) {
        current_start <- i
      }
    } else {
      if (!is.null(current_start)) {
        duration <- i - current_start
        if (duration >= min_duration) {
          fixation_periods[[length(fixation_periods) + 1]] <- c(current_start, i)
        }
        current_start <- NULL
      }
    }
  }
  
  # Check for fixation at end
  if (!is.null(current_start)) {
    duration <- n_samples - current_start + 1
    if (duration >= min_duration) {
      fixation_periods[[length(fixation_periods) + 1]] <- c(current_start, n_samples)
    }
  }
  
  return(fixation_periods)
}

# Calculate vergence
calculate_vergence <- function(data) {
  n_samples <- nrow(data)
  vergence_angles <- numeric(n_samples)
  
  for (i in 1:n_samples) {
    left_dir <- parse_vector_string(as.character(data[i, "LeftEyeDirection"]))
    right_dir <- parse_vector_string(as.character(data[i, "RightEyeDirection"]))
    
    if (length(left_dir) != 3 || length(right_dir) != 3 || any(is.na(c(left_dir, right_dir)))) {
      vergence_angles[i] <- NA
    } else {
      vergence_angles[i] <- calculate_angular_diff(left_dir, right_dir)
    }
  }
  
  return(vergence_angles)
}

# Main analysis function
analyze_eye_tracking <- function(data, window_size = 30) {
  # Calculate stability metrics
  stability_results <- calculate_gaze_stability(data, window_size)
  
  # Calculate vergence
  vergence_angles <- calculate_vergence(data)
  
  # Detect fixations
  fixation_periods <- detect_fixations(stability_results$angular_differences)
  
  # Calculate comprehensive metrics
  metrics <- list(
    # Stability metrics
    mean_stability = mean(stability_results$stability_scores, na.rm = TRUE),
    std_stability = sd(stability_results$stability_scores, na.rm = TRUE),
    max_angular_difference = max(stability_results$angular_differences, na.rm = TRUE),
    
    # Fixation metrics
    number_of_fixations = length(fixation_periods),
    total_fixation_time = sum(sapply(fixation_periods, function(x) x[2] - x[1])),
    fixation_periods = fixation_periods,
    
    # Eye metrics
    mean_left_openness = mean(data$LeftEyeOpeness, na.rm = TRUE),
    mean_right_openness = mean(data$RightEyeOpeness, na.rm = TRUE),
    mean_left_pupil = mean(data$LeftEyePupilSize, na.rm = TRUE),
    mean_right_pupil = mean(data$RightEyePupilSize, na.rm = TRUE),
    
    # Vergence metrics
    mean_vergence = mean(vergence_angles, na.rm = TRUE),
    std_vergence = sd(vergence_angles, na.rm = TRUE),
    
    # Raw data for plotting
    stability_scores = stability_results$stability_scores,
    angular_differences = stability_results$angular_differences,
    vergence_angles = na.omit(vergence_angles)
  )
  
  return(metrics)
}
```

## Analyzing Gaze Data (Entire Sample)

```{r warning=FALSE, message=FALSE}

# Execute Analysis
results <- analyze_eye_tracking(data_gam)

# Cumulative plotting function
plot_eye_tracking_cumulative <- function(results) {
  # Create basic plots using base R
  par(mfrow = c(2, 2))
  
  # Stability scores over time
  plot(results$stability_scores, type = "l",
       main = "Study_01 Gaze Stability Over Time",
       xlab = "Frame", ylab = "Stability Score", ylim = range(results$stability_scores, na.rm = TRUE))
  
  # Angular differences
  plot(results$angular_differences, type = "l",
       main = "Study_01 Angular Differences Between Frames",
       xlab = "Frame", ylab = "Angular Difference (degrees)", ylim = range(results$angular_differences, na.rm = TRUE))
  
  # Vergence angles
  plot(results$vergence_angles, type = "l",
       main = "Study_01 Vergence Angles Over Time",
       xlab = "Frame", ylab = "Vergence Angle (degrees)", ylim = range(results$vergence_angles, na.rm = TRUE))
  
  # Reset plot layout
  par(mfrow = c(1, 1))
}

# Plot Results
plot_eye_tracking_cumulative(results)
summary(results$stability_scores)
summary(results$angular_differences)
summary(results$vergence_angles)

```
## Gaze Analysis on Subject-by-Subject Basis

```{r warning=FALSE, message=FALSE}
# Split the data into a list of dataframes by Id
subject_data <- split(data_gam, data_gam$Id)

# Analyze each subject separately
subject_results <- lapply(subject_data, function(subject_df) {
  analyze_eye_tracking(subject_df)
})

# Function to plot eye tracking on subject-by-subject basis
plot_eye_tracking_subject <- function(results, subject_id) {
  par(mfrow = c(2, 2))
  
  # Stability scores over time
  plot(results$stability_scores, type = "l",
       main = paste("Study_01 Subject", subject_id, "- Gaze Stability"),
       xlab = "Frame", ylab = "Stability Score", ylim = range(results$stability_scores, na.rm = TRUE))
  
  #Angular differences
  plot(results$angular_differences, type = "l",
       main = paste("Study_01 Subject", subject_id, "- Angular Differences"),
       xlab = "Frame", ylab = "Angular Difference (degrees)", ylim = range(results$angular_differences, na.rm = TRUE))
  
  # Vergence angles
  plot(results$vergence_angles, type = "l",
      main = paste("Study_01 Subject", subject_id, "- Vergence Angles"),
       xlab = "Frame", ylab = "Vergence Angle (degrees)", ylim = range(results$vergence_angles, na.rm = TRUE))
  
  par(mfrow = c(1, 1))
}

# Create a directory to save plots (if it doesn't exist)
#dir.create("pre_study_subject_plots", showWarnings = FALSE)

# Generate and save plots for each subject
#lapply(names(subject_results), function(subject_id) {
  # Save plots to PNG files
  #png(paste0("subject_plots/subject_", subject_id, "_plots.png"), width = 800, height = 600)
  #plot_eye_tracking(subject_results[[subject_id]], subject_id)
  #dev.off()
#})

# Extract summary metrics for all subjects
summary_metrics <- do.call(rbind, lapply(names(subject_results), function(subject_id) {
  metrics <- subject_results[[subject_id]]
  data.frame(
    Subject = subject_id,
    Mean_Stability = metrics$mean_stability,
    Std_Stability = metrics$std_stability,
    Number_Fixations = metrics$number_of_fixations,
    Mean_Vergence = metrics$mean_vergence,
    Std_Vergence = metrics$std_vergence
  )
}))

na.omit(subject_results[1])


# Plot Results
plot_eye_tracking_subject(subject_results[[1]], 1)
plot_eye_tracking_subject(subject_results[[2]], 2)
plot_eye_tracking_subject(subject_results[[3]], 3)
plot_eye_tracking_subject(subject_results[[4]], 4)
plot_eye_tracking_subject(subject_results[[5]], 5)
plot_eye_tracking_subject(subject_results[[6]], 6)
plot_eye_tracking_subject(subject_results[[7]], 7)
plot_eye_tracking_subject(subject_results[[8]], 8)
plot_eye_tracking_subject(subject_results[[9]], 9)
plot_eye_tracking_subject(subject_results[[10]], 10)
plot_eye_tracking_subject(subject_results[[11]], 11)
plot_eye_tracking_subject(subject_results[[12]], 12)
plot_eye_tracking_subject(subject_results[[13]], 13)
plot_eye_tracking_subject(subject_results[[14]], 14)
plot_eye_tracking_subject(subject_results[[15]], 15)
plot_eye_tracking_subject(subject_results[[16]], 16)
plot_eye_tracking_subject(subject_results[[17]], 17)
plot_eye_tracking_subject(subject_results[[18]], 18)
plot_eye_tracking_subject(subject_results[[19]], 19)
plot_eye_tracking_subject(subject_results[[20]], 20)
plot_eye_tracking_subject(subject_results[[21]], 21)
plot_eye_tracking_subject(subject_results[[22]], 22)
plot_eye_tracking_subject(subject_results[[23]], 23)
plot_eye_tracking_subject(subject_results[[24]], 24)
plot_eye_tracking_subject(subject_results[[25]], 25)
plot_eye_tracking_subject(subject_results[[26]], 26)
plot_eye_tracking_subject(subject_results[[27]], 27)
# View the summary
print(summary_metrics)

```
## Gaze Analysis on Subject-by-Subject Basis

```{r warning=FALSE, message=FALSE}
# Add Condition column to full dataset
subject_condition_data <- data_gam #%>%
#  mutate(
  #  Condition = case_when(
  #    Phase == "Video" & Condition == "Main_Mono_Normal_HighLayerStrong" ~ "Normal_Mono",
 #     Phase == "Video" & Condition == "Main_Mono_Hallucinations_HighLayerStrong" ~ "Hallucination_Mono",
  #    Phase == "Video" & Condition == "Main_Stereo_Normal_HighLayerStrong" ~ "Normal_Stereo",
 #     Phase == "Video" & Condition == "Main_Stereo_Hallucinations_HighLayerStrong" ~ "Hallucination_Stereo",
 #     TRUE ~ NA_character_
 #   )
 # )

# Split data into nested structure: Subject -> Condition
subject_condition_data <- subject_condition_data %>%
  filter(Phase == "Video") %>%       # Keep only rows with valid conditions
  split(list(.$Id, .$Condition)) %>%  # Split into Subject-Condition groups
  lapply(function(x) {                # Remove grouping columns
    x %>% select(-Id, -Condition)
  })

# Analyze all subject-condition combinations
subject_condition_results <- lapply(
  names(subject_condition_data),
  function(name) {
    parts <- strsplit(name, "\\.")[[1]]
    list(
      subject_id = parts[1],
      condition = parts[2],
      results = analyze_eye_tracking(subject_condition_data[[name]])
    )
  }
)

plot_condition_comparison <- function(subject_id) {
  # Filter results for the target subject
  subject_results <- subject_condition_results[
    sapply(subject_condition_results, function(x) x$subject_id == subject_id)
  ]
  
  # Set up plot layout
  par(mfrow = c(3, 4), mar = c(4, 4, 2, 1))  # 3 metrics x 4 conditions
  
  # Plot each condition
  lapply(subject_results, function(sc) {
    # Stability Scores
    plot(sc$results$stability_scores, type = "l",
         main = paste("Subj", subject_id, sc$condition, "\nStability"),
         xlab = "", ylab = "Score",
         ylim = range(sc$results$stability_scores, na.rm = TRUE))
    
    # Angular Differences
    plot(sc$results$angular_differences, type = "l",
         main = paste("Subj", subject_id, sc$condition, "\nAngular Diff"),
         xlab = "", ylab = "Degrees",
         ylim = range(sc$results$angular_differences, na.rm = TRUE))
    
    # Vergence Angles
    plot(sc$results$vergence_angles, type = "l",
         main = paste("Subj", subject_id, sc$condition, "\nVergence"),
         xlab = "", ylab = "Degrees",
         ylim = range(sc$results$vergence_angles, na.rm = TRUE))
  })
  
  par(mfrow = c(1, 1))  # Reset layout
}

# Generate comparison summary
condition_summary <- do.call(rbind, lapply(subject_condition_results, function(sc) {
  data.frame(
    Subject = sc$subject_id,
    Condition = sc$condition,
    Mean_Stability = sc$results$mean_stability,
    Std_Stability = sc$results$std_stability,
    Mean_Vergence = sc$results$mean_vergence,
    Number_Fixations = sc$results$number_of_fixations
  )
}))

# Example usage
plot_condition_comparison("1") 

# View summary
print(condition_summary)

```

## Statistical Testing of Gaze Analysis

```{r warning=FALSE, message=FALSE}
# Function to run all specified comparisons for a metric
run_all_comparisons <- function(condition_summary, metric, comparisons_list) {
  results <- list()
  
  # Loop through each comparison pair in the list
  for (pair in comparisons_list) {
    condition1 <- pair[1]
    condition2 <- pair[2]
    
    tryCatch({
      result <- perform_paired_ttest(
        condition_summary = condition_summary,
        metric = metric,
        condition1 = condition1,
        condition2 = condition2
      )
      
      # Generate plot title
      plot_title <- paste(
        "Comparison:", metric, "\n",
        condition1, "vs", condition2
      )
      
      # Store results
      comparison_name <- paste(condition1, "vs", condition2, sep = "_")
      results[[comparison_name]] <- list(
        test_result = result,
        plot = plot_paired_gaze_comparison(result, plot_title)
      )
    }, error = function(e) {
      message("Skipping ", condition1, " vs ", condition2, " for ", metric, ": ", e$message)
    })
  }
  return(results)
}

# Main analysis workflow
analyze_all_metrics <- function(condition_summary, comparisons_list) {
  # Define metrics to analyze
  metrics <- c("Mean_Stability", "Mean_Vergence", "Number_Fixations")
  
  # Run analysis for each metric
  all_results <- list()
  for (metric in metrics) {
    all_results[[metric]] <- run_all_comparisons(
      condition_summary = condition_summary,
      metric = metric,
      comparisons_list = comparisons_list
    )
  }
  return(all_results)
}

# Execute the analysis
final_results <- analyze_all_metrics(condition_summary, comparisons)

# Example of accessing results:
# Print all stability plots
for (pair_result in final_results$Mean_Stability) {
  print(pair_result$plot)
}

# Get p-values for all vergence comparisons
vergence_results <- sapply(final_results$Mean_Vergence, function(x) {
  c(
    p.value = x$test_result$test_result$p.value,
    cohens.d = x$test_result$effect_size,
    n = x$test_result$n_pairs
  )
})

# Compile all results into a dataframe
results_df <- map_df(names(final_results), function(metric) {
  map_df(names(final_results[[metric]]), function(comparison) {
    res <- final_results[[metric]][[comparison]]$test_result
    data.frame(
      Metric = metric,
      Comparison = comparison,
      Mean_Difference = res$test_result$estimate,
      t_value = res$test_result$statistic,
      df = res$test_result$parameter,
      p_value = res$test_result$p.value,
      Cohen_d = res$effect_size,
      n_pairs = res$n_pairs,
      stringsAsFactors = FALSE
    )
  })
})

# Print the results dataframe
print(results_df)
```
