---
title: "Study 01 - Subjects Combined Questions-Physiological Analysis"
author:    |
    | Colin Ayres  
    | c.ayres@student.uw.edu.pl
date: "June 2025"
output: html_document
chunk_output_type: console
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Packages

```{r warning=FALSE, message=FALSE}
# Load packages
library(tidyverse)
library(here)
library(ggplot2)
library(zoo)
library(dplyr)
library(patchwork)
library(mgcv)
library(ggpubr)
library(rstatix)
library(readr)
library(gridExtra)
library(rmcorr) 
library(lme4)
library(lmerTest)
library(broom)
library(purrr)
library(reticulate)
library(stringr)

# Set working directory
mds_dir <- here::here()

```
## Load Physiological + Questions Data (From combined_heart.Rmd)

```{r warning=FALSE, message=FALSE}
pqm <- read.csv("study_01_processed_questions_merged.csv")

```

## Plotting Repeated Measures Correlation - ASC Score vs. HRV and HR Entropy

```{r warning=FALSE, message=FALSE}
# Group by Id and VideoName to calculate mean ASC_Score and heart_rate_variability
pqm_combined <- pqm %>%
  group_by(Id, VideoName) %>%
  summarise(
    ASC_Score = mean(ASC_Score, na.rm = TRUE),
    heart_rate_variability = mean(heart_rate_variability, na.rm = TRUE),
    HR_entropy = mean(HR_entropy)
  ) %>%
  ungroup()

# ASC vs HRV (Mono and Stereo ave)
ggplot(pqm_combined, aes(x = heart_rate_variability, y = ASC_Score, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Heart Rate Variability",
    y = "ASC Score",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# ASC vs HRV 
ggplot(pqm, aes(x = heart_rate_variability, y = ASC_Score, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Heart Rate Variability",
    y = "ASC Score",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Compute repeated measures correlation using combined mono/stereo scores
rmcorr_result_combined <- rmcorr(participant = Id, measure1 = ASC_Score, measure2 = heart_rate_variability, dataset = pqm_combined)

# View the results
print(rmcorr_result_combined)

# Compute repeated measures correlation without combined mono/stereo scores (4 points per subject)
rmcorr_result <- rmcorr(participant = Id, measure1 = ASC_Score, measure2 = heart_rate_variability, dataset = pqm)

# View the results
print(rmcorr_result)

# ASC vs HR Entropy 
ggplot(pqm, aes(x = HR_entropy, y = ASC_Score, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Heart Rate Entropy",
    y = "ASC Score",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")


# Compute repeated measures correlation without combined mono/stereo scores (7 points per subject)
rmcorr_result <- rmcorr(participant = Id, measure1 = ASC_Score, measure2 = HR_entropy, dataset = pqm)

# View the results
print(rmcorr_result)

```
## Do Answers to Questions 1 - 3 correlate with ASC Score?

```{r warning=FALSE, message=FALSE}
# Compute repeated measures correlation without combined mono/stereo scores (4 points per subject)
rmcorr_result_Q1 <- rmcorr(participant = Id, measure1 = ASC_Score, measure2 = Q1Answer, dataset = pqm)
rmcorr_result_Q2 <- rmcorr(participant = Id, measure1 = ASC_Score, measure2 = Q2Answer, dataset = pqm)
rmcorr_result_Q3 <- rmcorr(participant = Id, measure1 = ASC_Score, measure2 = Q3Answer, dataset = pqm)


# Q1 vs ASC
ggplot(pqm, aes(x = Q1Answer, y = ASC_Score, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Question 1 Answer",
    y = "ASC Score",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = Inf, hjust = -0.1, vjust = 1.1, size = 5, color = "black",
           label = sprintf("p = %.3e\nr = %.3f", rmcorr_result_Q1$p, rmcorr_result_Q1$r),
           family = "mono") 

# Q2 vs ASC
ggplot(pqm, aes(x = Q2Answer, y = ASC_Score, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Question 2 Answer",
    y = "ASC Score",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = Inf, hjust = -0.1, vjust = 1.1, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_Q2$p, rmcorr_result_Q2$r),
           family = "mono") 


# Q3 vs ASC
ggplot(pqm, aes(x = Q3Answer, y = ASC_Score, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Question 3 Answer",
    y = "ASC Score",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = Inf, hjust = -0.1, vjust = 1.1, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_Q3$p, rmcorr_result_Q3$r),
           family = "mono") 


print(rmcorr_result_Q1)

print(rmcorr_result_Q2)

print(rmcorr_result_Q3)
```
## Linear Regression Model for Questions 1/2/3 vs. ASC Score

```{r warning=FALSE, message=FALSE}
pqm$Condition_Order <- as.factor(pqm$Condition_Order)

analyze_and_plot <- function(response_var) {
# Fit the model
formula <- as.formula(paste("ASC_Score ~ ", response_var, " + Condition_Order + (1 | Id)"))
model <- lmer(formula, data = pqm)
print(paste("ASC ~", response_var, "ANOVA"))
print(summary(model))
print(anova(model))
print("-------------------------------")
  
# Plot the relationship
p1 <- ggplot(pqm, aes(x = .data[[response_var]], y = ASC_Score, color = Condition_Order)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = paste("ASC ~", response_var, "by Condition Type"), x = response_var, y = "ASC Score") +
  theme_minimal()
  
# Create a data frame with fitted values and residuals
resid_data <- data.frame(
  Fitted = fitted(model),
  Residuals = resid(model),
  Condition = pqm$Condition_Order
)

# Residuals vs. Fitted
p2 <- ggplot(resid_data, aes(x = Fitted, y = Residuals, color = Condition)) +
  geom_point(alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = paste("Residuals vs. Fitted ASC Score ~ ", response_var), x = paste("ASC Score Fitted Values"), y = paste("ASC Score Residuals")) +
  theme_minimal()
  
# QQ-plot
p3 <- ggplot(resid_data, aes(sample = Residuals)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = paste("QQ-Plot of Residuals for ASC Score ~ ", response_var)) +
  theme_minimal()

print(p1)
print(p2)
print(p3)
}

analyze_and_plot("Q1Answer")
analyze_and_plot("Q2Answer")
analyze_and_plot("Q3Answer")
```

## Linear Regression Model Between Q1-3 and ASC Score (Condition-Specific)

```{r warning=FALSE, message=FALSE}

run_lm_by_condition <- function(data, response_var) {
  response_var <- enquo(response_var)

  model_summaries <- data %>%
    group_by(Condition_Order) %>%
    nest() %>%
    mutate(
      model = map(data, ~ lm(as.formula(paste("ASC_Score ~", quo_name(response_var))), data = .)),
      tidy_model = map(model, broom::tidy)
    ) %>%
    unnest(tidy_model)

  r_values <- data %>%
    group_by(Condition_Order) %>%
    summarise(r = cor(ASC_Score, !!response_var, use = "complete.obs")) %>%
    mutate(r_label = paste0("r = ", round(r, 2)))

  pvals <- model_summaries %>% 
    filter(term == quo_name(response_var)) %>% 
    select(Condition_Order, p.value) %>% 
    mutate(p_label = paste0("p = ", signif(p.value, 2)))

  ggplot(data, aes_string(x = quo_name(response_var), y = "ASC_Score")) +
    geom_point(alpha = 0.7) +
    geom_smooth(method = "lm", se = FALSE, color = "blue") +
    facet_wrap(~ Condition_Order) +
    geom_text(data = pvals, aes(x = -Inf, y = Inf, label = p_label), hjust = -1.1, vjust = 2.4, size = 4, inherit.aes = FALSE) +
    geom_text(data = r_values, aes(x = -Inf, y = Inf, label = r_label), hjust = -1.4, vjust = 4.1, size = 4, inherit.aes = FALSE) +
    labs(title = paste("Regression of", quo_name(response_var), "on ASC_Score by Condition"),
         x = quo_name(response_var),
         y = "ASC_Score") +
    theme_minimal()
}

# Execution:
run_lm_by_condition(pqm, Q1Answer)
run_lm_by_condition(pqm, Q2Answer)
run_lm_by_condition(pqm, Q3Answer)

```


## Do Q1/Q2/Q3/ASC-Scores correlate with Drop in HRV (or Mean HR) from Baseline?

```{r warning=FALSE, message=FALSE}
# Process baseline data (remove outliers, calculate Mean HR and HRV)
data_base_processed_no_outliers <- filtered_baseline %>%
  group_by(Id) %>%
  mutate(
    # HRV (R-R) calculations
    avg_rr = mean(`Clean R-R`, na.rm = TRUE),
    lower_threshold_rr = 0.6 * avg_rr,
    upper_threshold_rr = 1.4 * avg_rr,
    is_outlier_rr = (`Clean R-R` < lower_threshold_rr) | (`Clean R-R` > upper_threshold_rr),
    Clean_RR_Filtered = ifelse(is_outlier_rr, NA, `Clean R-R`),
    
    # HR calculations
    avg_hr = mean(`Heart Rate`, na.rm = TRUE),
    lower_threshold_hr = 0.6 * avg_hr,
    upper_threshold_hr = 1.4 * avg_hr,
    is_outlier_hr = (`Heart Rate` < lower_threshold_hr) | (`Heart Rate` > upper_threshold_hr),
    Heart_Rate_Filtered = ifelse(is_outlier_hr, NA, `Heart Rate`)
  ) %>%
  ungroup()

baseline_results <- data_base_processed_no_outliers %>%
  filter(Id %in% 1:14) %>%
  group_by(Id) %>%
  arrange(Time) %>%
  summarise(
    mean_heart_rate = mean(`Heart Rate`, na.rm = TRUE),
    
    # Improved HRV calculation
    heart_rate_variability = {
      clean_rr <- na.omit(`Clean R-R`)
      if (length(clean_rr) >= 2) {
        sqrt(mean(diff(clean_rr)^2))  # RMSSD from consecutive non-NA values
      } else {
        NA_real_
      }
    },
    .groups = "drop"
  ) %>%
  mutate(
    mean_heart_rate = ifelse(is.nan(mean_heart_rate), NA, mean_heart_rate)
  )

# Add baseline differences for mean heart rate and heart rate variability
baseline_differences <- pqm %>%
  mutate(Id = as.integer(Id)) %>%
  left_join(
    baseline_results %>% 
      select(Id, baseline_mean_hr = mean_heart_rate, baseline_hrv = heart_rate_variability), 
    by = "Id"
  ) %>%
  mutate(
    baseline_difference_mean_hr = mean_heart_rate - baseline_mean_hr,
    baseline_difference_hrv = heart_rate_variability - baseline_hrv
  ) %>%
  select(-baseline_mean_hr, -baseline_hrv)  # Remove intermediate columns

# Compute repeated measures correlation without combined mono/stereo scores (4 points per subject)
rmcorr_result_BaselineHRV_Q1 <- rmcorr(participant = Id, measure1 = baseline_difference_hrv, measure2 = Q1Answer, dataset = baseline_differences)
rmcorr_result_BaselineHRV_Q2 <- rmcorr(participant = Id, measure1 = baseline_difference_hrv, measure2 = Q2Answer, dataset = baseline_differences)
rmcorr_result_BaselineHRV_Q3 <- rmcorr(participant = Id, measure1 = baseline_difference_hrv, measure2 = Q3Answer, dataset = baseline_differences)
rmcorr_result_BaselineHRV_ASC <- rmcorr(participant = Id, measure1 = baseline_difference_hrv, measure2 = ASC_Score, dataset = baseline_differences)


# Q1 vs HRV Difference from Baseline
ggplot(baseline_differences, aes(x = Q1Answer, y = baseline_difference_hrv, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Question 1 Answer",
    y = "HRV Deviation from Baseline",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -1.2, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_BaselineHRV_Q1$p, rmcorr_result_BaselineHRV_Q1$r),
           family = "mono") 

# Q2 vs HRV Difference from Baseline
ggplot(baseline_differences, aes(x = Q2Answer, y = baseline_difference_hrv, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Question 2 Answer",
    y = "HRV Deviation from Baseline",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -1.2, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_BaselineHRV_Q2$p, rmcorr_result_BaselineHRV_Q2$r),
           family = "mono")  

# Q3 vs HRV Difference from Baseline
ggplot(baseline_differences, aes(x = Q3Answer, y = baseline_difference_hrv, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Question 3 Answer",
    y = "HRV Deviation from Baseline",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -1.2, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_BaselineHRV_Q3$p, rmcorr_result_BaselineHRV_Q3$r),
           family = "mono") 

# ASC vs HRV Difference from Baseline
ggplot(baseline_differences, aes(x = ASC_Score, y = baseline_difference_hrv, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "ASC Score",
    y = "HRV Deviation from Baseline",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -1.2, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_BaselineHRV_ASC$p, rmcorr_result_BaselineHRV_ASC$r),
           family = "mono") 


print(rmcorr_result_BaselineHRV_Q1)
print(rmcorr_result_BaselineHRV_Q2)
print(rmcorr_result_BaselineHRV_Q3)
print(rmcorr_result_BaselineHRV_ASC)

# Compute repeated measures correlation without combined mono/stereo scores (4 points per subject)
rmcorr_result_BaselineMeanHR_Q1 <- rmcorr(participant = Id, measure1 = baseline_difference_mean_hr, measure2 = Q1Answer, dataset = baseline_differences)
rmcorr_result_BaselineMeanHR_Q2 <- rmcorr(participant = Id, measure1 = baseline_difference_mean_hr, measure2 = Q2Answer, dataset = baseline_differences)
rmcorr_result_BaselineMeanHR_Q3 <- rmcorr(participant = Id, measure1 = baseline_difference_mean_hr, measure2 = Q3Answer, dataset = baseline_differences)
rmcorr_result_BaselineMeanHR_ASC <- rmcorr(participant = Id, measure1 = baseline_difference_mean_hr, measure2 = ASC_Score, dataset = baseline_differences)


# Q1 vs HRV Difference from Baseline
ggplot(baseline_differences, aes(x = Q1Answer, y = baseline_difference_mean_hr, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Question 1 Answer",
    y = "Mean HR Deviation from Baseline",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -1.2, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_BaselineMeanHR_Q1$p, rmcorr_result_BaselineMeanHR_Q1$r),
           family = "mono") 

# Q2 vs HRV Difference from Baseline
ggplot(baseline_differences, aes(x = Q2Answer, y = baseline_difference_mean_hr, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Question 2 Answer",
    y = "Mean HR Deviation from Baseline",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -1.2, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_BaselineMeanHR_Q2$p, rmcorr_result_BaselineMeanHR_Q2$r),
           family = "mono")  

# Q3 vs HRV Difference from Baseline
ggplot(baseline_differences, aes(x = Q3Answer, y = baseline_difference_mean_hr, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Question 3 Answer",
    y = "Mean HR Deviation from Baseline",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -1.2, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_BaselineMeanHR_Q3$p, rmcorr_result_BaselineMeanHR_Q3$r),
           family = "mono") 

# ASC vs HRV Difference from Baseline
ggplot(baseline_differences, aes(x = ASC_Score, y = baseline_difference_mean_hr, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "ASC Score",
    y = "Mean HR Deviation from Baseline",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -1.2, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_BaselineMeanHR_ASC$p, rmcorr_result_BaselineMeanHR_ASC$r),
           family = "mono") 


print(rmcorr_result_BaselineMeanHR_Q1)
print(rmcorr_result_BaselineMeanHR_Q2)
print(rmcorr_result_BaselineMeanHR_Q3)
print(rmcorr_result_BaselineMeanHR_ASC)

```

## Mean Pupil Size Difference from Baseline vs. Q1/Q2/Q3/ASC Answers

```{r warning=FALSE, message=FALSE}
# Define file paths
file_paths <- list.files(path = "Study_01", pattern = "_DeviceData.csv", full.names = TRUE)

# Import CSV files and combine them into a single data frame
device_combined <- bind_rows(lapply(file_paths, function(file) {
  read_csv2(file, 
            col_types = cols(Id = col_integer(), SegmentIndex = col_integer(), QuestionIndex = col_double(), LeftEyePupilSize = col_double(), RightEyePupilSize = col_double(), .default = col_guess()),
            show_col_types = FALSE)
}))

condensed_device_combined <- device_combined %>%
  mutate(
    Condition = as.factor(VideoMode),
    VideoMode = case_when(
      grepl("Main_Stereo", VideoName) ~ "Stereo",
      grepl("Main_Mono", VideoName) ~ "Mono",
      TRUE ~ as.character(VideoMode)
    ),
    VideoName = as.factor(VideoName)
  ) %>%
  group_by(Id) %>%
  mutate(Condition_Order = paste(VideoMode, Condition, sep = "_")) %>%
  unnest_longer(Condition_Order)

mean_pupil_size_df <- condensed_device_combined %>%
  filter(Condition_Order != "NA_NA") %>%
  group_by(Id, Condition_Order) %>%
  summarise(
    MeanAveragePupilSize = mean((LeftEyePupilSize + RightEyePupilSize) / 2, na.rm = TRUE),
    .groups = 'drop'
  )

# Define file paths
file_paths <- list.files(path = "Study_01", pattern = "_Baseline.csv", full.names = TRUE)

# Import CSV files and combine them into a single data frame
baseline_combined <- bind_rows(lapply(file_paths, read_csv2))

mean_pupil_size_df_baseline <- baseline_combined %>%
  group_by(Id) %>%
  summarise(
    MeanAveragePupilSizeBaseline = mean((LeftEyePupilSize + RightEyePupilSize) / 2, na.rm = TRUE),
    .groups = 'drop'
  )

final_pupil_size_df <- mean_pupil_size_df %>%
  left_join(mean_pupil_size_df_baseline %>% select(Id, MeanAveragePupilSizeBaseline), by = "Id") %>%
  mutate(MeanPupilSize_Baseline_Diff = MeanAveragePupilSize - MeanAveragePupilSizeBaseline) %>%
  select(Id, Condition_Order, MeanPupilSize_Baseline_Diff)

baseline_differences <- baseline_differences %>%
  left_join(final_pupil_size_df %>% select(Id, Condition_Order, MeanPupilSize_Baseline_Diff), by = c("Id", "Condition_Order"))

# Compute repeated measures correlation without combined mono/stereo scores (7 points per subject)
rmcorr_result_BaselineMeanPupil_Q1 <- rmcorr(participant = Id, measure1 = MeanPupilSize_Baseline_Diff, measure2 = Q1Answer, dataset = baseline_differences)
rmcorr_result_BaselineMeanPupil_Q2 <- rmcorr(participant = Id, measure1 = MeanPupilSize_Baseline_Diff, measure2 = Q2Answer, dataset = baseline_differences)
rmcorr_result_BaselineMeanPupil_Q3 <- rmcorr(participant = Id, measure1 = MeanPupilSize_Baseline_Diff, measure2 = Q3Answer, dataset = baseline_differences)
rmcorr_result_BaselineMeanPupil_ASC <- rmcorr(participant = Id, measure1 = MeanPupilSize_Baseline_Diff, measure2 = ASC_Score, dataset = baseline_differences)

# Q1 vs HRV Difference from Baseline
ggplot(baseline_differences, aes(x = Q1Answer, y = MeanPupilSize_Baseline_Diff, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Question 1 Answer",
    y = "Mean Pupil Size Deviation from Baseline",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -1.2, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_BaselineMeanPupil_Q1$p, rmcorr_result_BaselineMeanPupil_Q1$r),
           family = "mono") 

# Q2 vs HRV Difference from Baseline
ggplot(baseline_differences, aes(x = Q2Answer, y = MeanPupilSize_Baseline_Diff, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Question 2 Answer",
    y = "Mean Pupil Size Deviation from Baseline",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -1.2, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_BaselineMeanPupil_Q2$p, rmcorr_result_BaselineMeanPupil_Q2$r),
           family = "mono")  

# Q3 vs HRV Difference from Baseline
ggplot(baseline_differences, aes(x = Q3Answer, y = MeanPupilSize_Baseline_Diff, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "Question 3 Answer",
    y = "Mean Pupil Size Deviation from Baseline",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -1.2, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_BaselineMeanPupil_Q3$p, rmcorr_result_BaselineMeanPupil_Q3$r),
           family = "mono") 

# ASC vs HRV Difference from Baseline
ggplot(baseline_differences, aes(x = ASC_Score, y = MeanPupilSize_Baseline_Diff, color = as.factor(Id), shape = VideoName)) +
  geom_point(size = 2) +  # Scatterplot points
  geom_smooth(aes(group = Id), method = "lm", se = FALSE, linetype = "solid") + 
  labs(
    x = "ASC Score",
    y = "Mean Pupil Size Deviation from Baseline",
    color = "Subject ID",
    shape = "Video Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -1.2, size = 5, color = "black",
           label = sprintf("p = %.3f\nr = %.3f", rmcorr_result_BaselineMeanPupil_ASC$p, rmcorr_result_BaselineMeanPupil_ASC$r),
           family = "mono") 


print(rmcorr_result_BaselineMeanPupil_Q1)
print(rmcorr_result_BaselineMeanPupil_Q2)
print(rmcorr_result_BaselineMeanPupil_Q3)
print(rmcorr_result_BaselineMeanPupil_ASC)

```
